Cycles of the logistic map

### 2013-06-05 doc/pc-ijbc.tex
modify mknsolv.py
add bmp.ma, bmp.c and bmp.py

% switch text to two-column
%\begin{multicols}{2}
\twocolumn

@@

%\begin{table}[h]
%  %\small
%  \tbl{
%  Smallest positive $r$ at the onset and bifurcation points
%  of the $n$-cycles of the logistic map.
%  }
%{
%\begin{tabular}{llll}
%%\begin{tabularx}{\linewidth}{
%%  >{\hsize=0.5\hsize\centering\arraybackslash}X
%%  >{\hsize=1.6\hsize}X
%%  >{\hsize=1.6\hsize}X
%%  >{\hsize=0.3\hsize\raggedright\arraybackslash}X
%%}
%\hline
%  $n^\dagger$
%& Onset$^\ddagger$
%& Bifurcation$^\ddagger$
%& \#$^*$ \\
%\hline
%$1$     & $1.0000000000_1$      &  $3.0000000000_1$       & 1   \\
%$2'$    & $3.0000000000_1$      &  $3.4494897428_1$       & 1   \\
%$3$     & $3.8284271247_1$      &  $3.8414990075_3$       & 1   \\
%$4$     & $3.9601018827_3$      &  $3.9607686524_6$       & 1   \\
%$4''$   & $3.4494897428_1$      &  $3.5440903596_6$       & 1   \\
%$5$     & $3.7381723753_{11}$   &  $3.7411207566_{15}$    & 3   \\
%$6$     & $3.6265531617_{20}$   &  $3.6303887000_{27}$    & 4   \\
%$6'$    & $3.8414990075_{3}$    &  $3.8476106612_{27}$    & 1   \\
%$7$     & $3.7016407642_{57}$   &  $3.7021549282_{63}$    & 9   \\
%$8$     & $3.6621089132_{108}$  &  $3.6624407072_{120}$   & 14  \\
%$8'$    & $3.9607686524_{6}$    &  $3.9610986335_{120}$   & 1   \\
%$8'''$  & $3.5440903596_{6}$    &  $3.5644072661_{120}$   & 1   \\
%$9$     & $3.6871968733_{240}$  &  $3.6872742105_{252}$   & 28  \\
%$10$    & $3.6052080669_{472}$  &  $3.6059169323_{495}$   & 48  \\
%$10'$   & $3.7411207566_{15}$   &  $3.7425706462_{495}$   & 3   \\
%$11$    & $3.6817160194_{1013}$ &  $3.6817266457_{1023}$  & 93  \\
%$12$    & $3.5820230011_{1959}$ &  $3.5828117795_{2010}$  & 165 \\
%$12'$   & $3.6303887000_{27}$   &  $3.6321857392_{2010}$  & 4   \\
%$12''$  & $3.8476106612_{27}$   &  $3.8490363152_{2010}$  & 1   \\
%$13$    & $3.6797024578_{4083}$ &  $3.6797038498_{4095}$  & 315 \\
%\hline
%\multicolumn{4}{p{\linewidth}}{
%$^\dagger$
%  $\,'$, $\,''$, or $\,'''$ means
%    that the cycle is undergoing
%    the first, second, or third successive period-doubling, respectively.
%} \\
%\multicolumn{4}{p{\linewidth}}{
%$^\ddagger$
%  The subscripts are the degrees of the corresponding minimal polynomial
%    of $R = r(r-2)/4$.
%} \\
%\multicolumn{4}{p{\linewidth}}{
%$^*$
%  The number of similar cycles.
%} \\
%\hline
%%\end{tabularx}
%\end{tabular}
%\label{tab:rval}
%}
%\end{table}

@@

We coded the algorithm into a Mathematica program,
and computed the polynomials for $n$ up to 14.
%
As shown in \reffig{logtime}, the algorithm
  outperformed the resultant \cite{burm} and
  Gr\"obner basis \cite{kk1} methods,
  by at least two orders of magnitudes.
%
This is unsurprising,
  because the resultant of $f^n(x)-x$ and $\frac{d}{dx}f^n(x)-\lambda$
  is the determinant of the Sylvester matrix, which has
  $2^{n+1}-1$ rows and columns
  [as $\deg_x f^n(x) = 2^n$],
  while the matrix in \refeq{xr}
  is of only $\NB \approx 2^n/n$ [cf. \refeq{necklace}] rows and columns.
%Thus, the current algorithm is expected to be faster.

@@

%
The exact polynomials of these maps are generally too large to print on paper,
  e.g., the bifurcation polynomial of the 14-cycle
    takes about 27 megabytes.
%
The polynomials and programs are saved on the website
http://logperiod.appspot.com.
%
%The representative $r$ values are listed in \reftab{rval}.

%Compared with other approaches \cite{hitzl, gordon, bailey2, kk1, hao, hao2}
%it appeared to be simpler and more transparent,
%as it only requires the computation of a determinant.

>>>>

% switch text to two-column
%\begin{multicols}{2}
%\twocolumn

@@

%\begin{table}[h]
%  %\small
%  \tbl{
%  Smallest positive $r$ at the onset and bifurcation points
%  of the $n$-cycles of the logistic map.
%  }
%{
%\begin{tabular}{llll}
%%\begin{tabularx}{\linewidth}{
%%  >{\hsize=0.5\hsize\centering\arraybackslash}X
%%  >{\hsize=1.6\hsize}X
%%  >{\hsize=1.6\hsize}X
%%  >{\hsize=0.3\hsize\raggedright\arraybackslash}X
%%}
%\hline
%  $n^\dagger$
%& Onset$^\ddagger$
%& Bifurcation$^\ddagger$
%& \#$^*$ \\
%\hline
%$1$     & $1.0000000000_1$      &  $3.0000000000_1$       & 1   \\
%$2'$    & $3.0000000000_1$      &  $3.4494897428_1$       & 1   \\
%$3$     & $3.8284271247_1$      &  $3.8414990075_3$       & 1   \\
%$4$     & $3.9601018827_3$      &  $3.9607686524_6$       & 1   \\
%$4''$   & $3.4494897428_1$      &  $3.5440903596_6$       & 1   \\
%$5$     & $3.7381723753_{11}$   &  $3.7411207566_{15}$    & 3   \\
%$6$     & $3.6265531617_{20}$   &  $3.6303887000_{27}$    & 4   \\
%$6'$    & $3.8414990075_{3}$    &  $3.8476106612_{27}$    & 1   \\
%$7$     & $3.7016407642_{57}$   &  $3.7021549282_{63}$    & 9   \\
%$8$     & $3.6621089132_{108}$  &  $3.6624407072_{120}$   & 14  \\
%$8'$    & $3.9607686524_{6}$    &  $3.9610986335_{120}$   & 1   \\
%$8'''$  & $3.5440903596_{6}$    &  $3.5644072661_{120}$   & 1   \\
%$9$     & $3.6871968733_{240}$  &  $3.6872742105_{252}$   & 28  \\
%$10$    & $3.6052080669_{472}$  &  $3.6059169323_{495}$   & 48  \\
%$10'$   & $3.7411207566_{15}$   &  $3.7425706462_{495}$   & 3   \\
%$11$    & $3.6817160194_{1013}$ &  $3.6817266457_{1023}$  & 93  \\
%$12$    & $3.5820230011_{1959}$ &  $3.5828117795_{2010}$  & 165 \\
%$12'$   & $3.6303887000_{27}$   &  $3.6321857392_{2010}$  & 4   \\
%$12''$  & $3.8476106612_{27}$   &  $3.8490363152_{2010}$  & 1   \\
%$13$    & $3.6797024578_{4083}$ &  $3.6797038498_{4095}$  & 315 \\
%$14$    & $3.5972109838_{8052}$ &  $3.5973201609_{8127}$  & 576 \\
%$14'$   & $3.7021549282_{63}$   &  $3.7024116236_{8127}$  & 9   \\
%\hline
%\multicolumn{4}{p{\linewidth}}{
%$^\dagger$
%  $\,'$, $\,''$, or $\,'''$ means
%    that the cycle is undergoing
%    the first, second, or third successive period-doubling, respectively.
%} \\
%\multicolumn{4}{p{\linewidth}}{
%$^\ddagger$
%  The subscripts are the degrees of the corresponding minimal polynomial
%    of $R = r(r-2)/4$.
%} \\
%\multicolumn{4}{p{\linewidth}}{
%$^*$
%  The number of similar cycles.
%} \\
%\hline
%%\end{tabularx}
%\end{tabular}
%\label{tab:rval}
%}
%\end{table}

@@

\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[width=\linewidth]{T14b.png}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:T14b}
  Coefficients of the 14-cycle bifurcation polynomial in two columns (left to right).
  The $k$th line from the top is the coefficient before $(4R)^k$ in base-256.
  The digits are shown by the grayscale from white (0) to black (255).
  The leading bar before each line is the sign (black: $-$, white: $+$).
  }
\end{figure}

We coded the algorithm into a Mathematica program,
  and computed the polynomials for $n$ up to 14
  (data on the website http://logperiod.appspot.com).
%
The bifurcation polynomial of the 14-cycle in shown in \reffig{T14b}.
%
%The exact polynomials of these maps are generally too large to print on paper,
%  e.g., the bifurcation polynomial of the 14-cycle
%    takes about 27 megabytes.
%
%The representative $r$ values are listed in \reftab{rval}.

%Compared with other approaches \cite{hitzl, gordon, bailey2, kk1, hao, hao2}
%it appeared to be simpler and more transparent,
%as it only requires the computation of a determinant.
%
As shown in \reffig{logtime}, the algorithm
  outperformed the resultant \cite{burm} and
  Gr\"obner basis \cite{kk1} methods,
  by at least two orders of magnitudes.
%
This is unsurprising,
  because the resultant of $f^n(x)-x$ and $\frac{d}{dx}f^n(x)-\lambda$
  is the determinant of the Sylvester matrix, which has
  $2^{n+1}-1$ rows and columns
  [as $\deg_x f^n(x) = 2^n$],
  while the matrix in \refeq{xr}
  is of only $\NB \approx 2^n/n$ [cf. \refeq{necklace}] rows and columns.
%Thus, the current algorithm is expected to be faster.

@@
